---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo = TRUE}
options("scipen"=10)
#set high penalty for scientific number display
library("lubridate")
work <- read.csv("activity.csv")
work$date <- ymd(work$date)


```
## What is mean total number of steps taken per day?
```{r, echo = TRUE}
library("dplyr")
library("ggplot2")
ggplot(work, aes(x=date, y = steps)) + geom_histogram(binwidth=.5, stat="identity") + ylab("total number of steps")

work_tbl <- tbl_df(work)
by_date <- group_by(work_tbl, date)
total_steps_by_date <- summarise(by_date, total_steps = sum(steps))
total_steps_mean <- mean(total_steps_by_date$total_steps, na.rm=TRUE)
total_steps_median <- median(total_steps_by_date$total_steps, na.rm=TRUE)

```

the mean total number of steps taken per day is `r total_steps_mean`.

the median total number of steps taken per day is `r total_steps_median`.
 
## What is the average daily activity pattern?
```{r, echo = TRUE}
library("dplyr")
library("ggplot2")

by_interval <- group_by(work_tbl, interval)
mean_steps_by_interval <- summarise(by_interval, mean_steps = mean(steps, na.rm=TRUE))

ggplot(mean_steps_by_interval, aes(x= interval, y= mean_steps)) + geom_line()+ xlab("inverval (5 minutes per unit)")
# draw a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

arrange_mean <- arrange(mean_steps_by_interval, desc(mean_steps))
#
```
"`r arrange_mean[[1,1]]`"is the interval, on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values
```{r}
total_na <- sum(is.na(work$steps))
```
The total number of missing values in the dataset is `r total_na`.
```{r}
na_in_work <-
    work %>% 
        group_by(date) %>%
        mutate(steps= replace(steps, is.na(steps), mean(steps, na.rm=TRUE))) %>%
        #replace na with the average steps of that date.
        mutate(steps= replace(steps, is.nan(steps), 0))
        #replace NaN with 0

ggplot(na_in_work, aes(x=date, y = steps)) + geom_histogram(binwidth=.5, stat="identity") + ylab("total number of steps")


by_date <- group_by(na_in_work, date)
total_steps_by_date <- summarise(by_date, total_steps = sum(steps))
total_steps_mean_n <- mean(total_steps_by_date$total_steps)
total_steps_median_n <- median(total_steps_by_date$total_steps)
```
The mean total number of steps taken per day is `r total_steps_mean_n`

The median total number of steps taken per day is `r total_steps_median_n`

Yes, the mean and the median differ from the estimates from the first part of the assignment.

I try to use the mean of that day to fill the NA, but because the NA exist in whole-day way, I decide to fill them by using number zero, or I will get NaN.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday_work <-
    na_in_work %>%
        mutate(weekdayn = wday(date)) %>%
        #find out which week day
        mutate(weekday = ifelse(weekdayn == 1 | weekdayn == 7, "weekend", "weekday"))
        #if saturday or sunday, it's weekend, else it's weekday.


weekday_work_d <- filter(weekday_work, weekday == "weekday")
weekend_work_d <- filter(weekday_work, weekday == "weekend")

mean_steps_weekday_n <-
    weekday_work_d %>%
        group_by(interval) %>%
        summarise(mean_steps = mean(steps))

mean_steps_weekend_n <-
    weekend_work_d %>%
        group_by(interval) %>%
        summarise(mean_steps = mean(steps))


plot_weekday <- ggplot(mean_steps_weekday_n, aes(x= interval, y= mean_steps)) + geom_line()+ xlab("inverval (5 minutes per unit)")+labs(title = "weekday")
plot_weekend <- ggplot(mean_steps_weekend_n, aes(x= interval, y= mean_steps)) + geom_line()+ xlab("inverval (5 minutes per unit)") + labs(title = "weekend")

library(grid)
pushViewport(viewport(layout = grid.layout(2, 1)))
print(plot_weekday, vp = viewport(layout.pos.row = 2, layout.pos.col = 1))
print(plot_weekend, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
```

By seeing the plot, there are differences in activity patterns between weekdays and weekends
